Maximum Entropy Prediction Markets

### 2014-02-10 WriteUp/maxentpredmkts.tex
some more comments on budgets

\section{The Exponential Family Market Mechanism with Budgets} \label{sec:budgets}
\sindhu{\cite{Azoury01}}
In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.

@@

In this section, we assume that the traders have exponential family beliefs. The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
\subsection{Budget-limited Aggregation}
Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices. We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market. This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader. Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.

In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
Let $\alpha$ be the budget of a trader in the market. Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief. Let $C$ be the cost function. Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief. In this case, we want to budget-limit the trader's influence on the market state.

@@

First, we  show that this trade is feasible given the trader's budget:
\begin{theorem}
Let the current market state be given by $\theta$. Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$. The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
\end{theorem}
\begin{proof}
From the convexity of $C$, we have
\begin{eqnarray*}
C(\theta')&\leq&(1-\lambda)C(\theta)+\lambda C(\htheta)\\
\mbox{Thus, } C(\theta') - C(\theta) &\leq& \lambda\ (C(\htheta) - C(\theta))\\
\mbox{And, } C(\theta') - C(\theta) &\leq& \alpha 
 \end{eqnarray*}\end{proof}

>>>>

\section{The Exponential Family Market Mechanism with Budgets} \label{sec:budgets}
\sindhu{\cite{Azoury01}}
%In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.

@@

In this section, we assume that the traders have exponential family beliefs. The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$; the vector of sufficient statistics is assumed to be non-negative. \sindhu{is this ok?}.
\subsection{Budget-limited Aggregation}
Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices. We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market. This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader. Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.

In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. \sindhu{this assumption is also very strong but is needed}
Let $\alpha$ be the budget of a trader in the market. Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief. Let $C$ be the cost function. Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief. In this case, we want to budget-limit the trader's influence on the market state.

@@

First, we  show that this trade is feasible given the trader's budget:
\begin{lemma}
Let the current market state be given by $\theta$. Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$. The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
\end{lemma}
\begin{proof}
From the convexity of $C$, we have
\begin{eqnarray*}
C(\theta')&\leq&(1-\lambda)C(\theta)+\lambda C(\htheta)\\
\mbox{Thus, } C(\theta') - C(\theta) &\leq& \lambda\ (C(\htheta) - C(\theta))\\
\mbox{And, } C(\theta') - C(\theta) &\leq& \alpha 
 \end{eqnarray*}\end{proof}
